---
title: "Assignment_2"
author: "Manav-400451088"
date: "2023-02-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34)
library(dplyr)
library(readr)
library(magrittr)
library(knitr)
library(tidyverse)
```

```{r, echo=FALSE, eval=FALSE}
idrfunction <- function(x) {
  quan <- unname(quantile(x, probs = seq(0, 1, 1/10), na.rm = TRUE))
  ans <- quan[9] - quan[1]
  return(ans)
}


tidyfunc <- myx,na.rm = TRUE)
    idr_ <- idrfunction(myx)
   q1b <- c(sd=sd_,idr=idr_) %>%
    round(2)
}

idr_sd_infant_mortality <- gapminder %>%
  pull(infant_mortality)%>%
  tidyfunc()
idr_sd_fertility <- gapminder %>%
  pull(fertility)%>%
  tidyfunc()

final_ans <- c(fertlity=idr_sd_fertility,infant_mortality=idr_sd_infant_mortality)
final_ans
```
# Question 2

### The data I am analyzing is of temperatures throughout the months of September to January 2022 -> 2023
### The columns include:
* CHR1 - Will be removed due to no importance to the data
* CHR4 - Will be renamed into "temperature" because it shows the temperature
* SEP/22 - September 2022, will be lowered
* Oct/22 - October 2022, will be lowered
* nov/22 - November 2022, will be kept the same
* dEc/22 - December 2022, will be lowered
* JaN/23 - January 2023, will be lowered

# Iteration 1
## Phase 1
1. Read CSV file into R
    * Specify column types?
2. Change the column names to lowercase
3. Identify the uids
    * Check that they are unique
4. Check if the data is in Tidy format
5. Sort tibble by uids.

# Phase 2
```{r,echo=TRUE,eval=TRUE}
###Read CSV files into R
yay <- read_csv("./as2.csv", col_types = "ccnnnnn", col_names = c("trash", "feels_like","sep_2022","oct_2022","nov_2022","dec_2022","jan_2023"))
###Making all the column names look better and removing the / since
###/ is a divider and will get annoying to work with columns with it

#"trash" seems to just be a random letter column
yay %>% count(trash) %>% filter(n>1) %>% slice(1:5) %>% kable()
#feels_like has a few problems such as "M", "MED" not being consistent with med
yay %>% count(feels_like) %>% filter(n>1) %>% slice(1:5) %>% kable()
yay %>%
  glimpse()
```
### Based on the results, "feels_like" looks like a good candiate for the uid
# Iteration 2
## Phase 1
1. Check for NA values in each columns
2. Look for outliers in numeric columns
3. Check the values in the character columns
4. Double check

## Phase 2
```{r,echo=TRUE,eval=TRUE}
###Based on the summary there are outliers and NA values throughout the data
summary(yay)

###Lets first get rid of the trash column
yay2 <- yay %>%
  select(-c(trash))

###Now if we display this we see that the first row is actually actually useless 
###since it doesn't correlate with the data being given.
yay2 
###Removing the first row and removing NA rows.
yay3 <- yay2 %>%
  drop_na() %>%
  slice(-1)
yay3
###Now if we look at feels_like column we see the rows are inconsistent.
yay3 %>% count(feels_like) %>% filter(n>1) %>% slice(1:5) %>% kable()
###So lets fix that
yay4 <- yay3 %>% 
  mutate(feels_like = str_replace(feels_like, "MED", "med"))%>%
  mutate(feels_like = str_replace(feels_like, "M", "med"))%>%
  mutate(feels_like = str_replace(feels_like, "medEdIUM", "med"))%>%
  mutate(feels_like = str_replace(feels_like, "mededium", "med"))
yay4 %>% count(feels_like) %>% filter(n>1) %>% slice(1:5) %>% kable()
###Now if we check the summary we see the NA values are gone and same with
###the trash column.
summary(yay4)
###Now lets gets rid of the outliers. By inspection we can see the last row
###of "yay4" is filled with outliers so we can simply remove the row.
yay_final <- yay4 %>%
  slice(-92)

###Now finally, we doublecheck for completeness

yay_final%>%pull(feels_like)%>% is.na()%>%any()
yay_final%>%pull(sep_2022)%>% is.na()%>%any()
yay_final%>%pull(oct_2022)%>% is.na()%>%any()
yay_final%>%pull(nov_2022)%>% is.na()%>%any()
yay_final%>%pull(dec_2022)%>% is.na()%>%any()
yay_final%>%pull(jan_2023)%>% is.na()%>%any()

###We check the summary once again.
###The summary looks good
summary(yay_final)

###Finally, we have the final and cleaned data
kable(yay_final)


```


