---
title: "DWP2"
author: "Manav-400451088"
date: "2023-03-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34)
library(tidyr)
library(tidyverse)
library(magrittr)
library(knitr)
opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```
## Iteration 1
### Phase 1
1. Read in CSV file
2. Make all columns into character
3. Convert column names to lowercase
4. Determine the uid's, get rid of index columns
5. Tidy the data
6. Reorder the columns so uid's are first
7. Sort by uid's
```{r}
#1,2
data <- read_csv("/Users/manavsharma/Desktop/Data Science/console.csv", col_types = "ccccccc") %>%
  #3
  rename_with(tolower)

data %>%
  glimpse()

```
Determine UID
```{r}
data %>%
  count(console)%>%
  filter(n>1)
```
Console is a UID
```{r}
data %<>% 
  select(!index)

data1 <- data %>%
  pivot_longer(!console,names_to = "region",values_to = "sales")

data1 %>%
  glimpse()
```
Now we sort by UID
```{r}
data1 %<>% arrange(console)
data1 %>%
  slice(1:5) %>%
  kable()

```
## Iteration 2
### Phase 1
1. Convert all columns into factor
2. Check for Ouliars and set to NA
3. Remove all NA's

```{r}
data1 %<>%
  mutate(console = as.factor(console), region = as.factor(region), sales = as.factor(sales))

data1 %>%
  summary()
```

Sales has 80 NA's. Lets remove them.
```{r}
data1 %<>%
  drop_na(sales)

data1 %$%
  is.na(sales)%>%
  any()
```
Now we have got rid of all NA's.

The DWP is now finished
```{r}
data1 %>%
  kable()
```


