---
title: "DWPSOLO2"
author: "Manav-400451088"
date: "2023-03-08"
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE, echo=FALSE, eval=TRUE}
library(tidyverse)
library(magrittr)
library(knitr)

set.seed(34)

opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

1. Read CSV file into `R`
    - Specify column types
2. Change the column names to lowercase
3. Identify the uid's
    - Check that they are unique
4. Check if the data is in `Tidy` format
5. Sort tibble by uids
```{r}
#1,2
tib <- read_csv("/Users/manavsharma/Desktop/Data Science/vgsales.csv", col_types = "ccccccccccc")
tib %<>%
  rename_with(tolower)
tib %>%
  glimpse()
```
Identifying the UID's
```{r}
tib %>%
  count(rank)%>%
  filter(n>1) #rank is unique

tib %>%
  count(genre)%>%
  filter(n>1) #genre is unique
```
Now that we have our UID's we will tidy the data.
```{r}
tib_ <- tib %<>%
  pivot_longer(!c(rank,name,platform,year,genre,publisher), values_to = "values", names_to = "sales")

tib_ %>%
  glimpse()
```
Now our data is tidy.
Now we sort by our UID's
```{r}
tib_ %<>%
  select(rank,genre,everything())
tib_ %>%
  glimpse()
```
# Iteration 2

## Phase 1
1. Convert all columns to factor
2. Look for outliers in numeric columns, make them into NA
3. Check for NA values, and then drop them all.

## Phase 2
```{r}
tib_ %<>%
  mutate(rank=as.factor(rank),genre = as.factor(genre),name = as.factor(name),platform = as.factor(platform),year = as.factor(year),publisher = as.factor(publisher),sales = as.factor(sales),values=as.factor(values))

tib_ %>%
  summary()
```
Rank is good
Genre is good
Name is good
Platform is good
Year is good
Publisher is good
Sales is good
Values is good

Now we check just incase
```{r}
tib_ %$%
  is.na(rank)%>%
  any()
tib_ %$%
  is.na(genre)%>%
  any()
tib_ %$%
  is.na(name)%>%
  any()
tib_ %$%
  is.na(platform)%>%
  any()
tib_ %$%
  is.na(year)%>%
  any()
tib_ %$%
  is.na(publisher)%>%
  any()
tib_ %$%
  is.na(year)%>%
  any()
tib_ %$%
  is.na(values)%>%
  any()
```
Alright we good now,
```{r}
tib_ %>%
  slice(1:50) %>%
  kable()
```

