---
title: "Midterm1DWP"
author: "Manav Sharma 400451088"
date: "2023-03-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(knitr)
opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, error = FALSE)
set.seed(34)
```
# Iteration 1

## Phase 1

1. Read CSV file into `R`
2. Specify column types
3. Change the column names to lowercase.
4. Identify the uid's.
5. Check that they are unique.
6. Check if the data is in `Tidy` format.
7. Sort tibble by uids.

## Phase 2
```{r}
#1 & 2
data <- read_csv("/Users/manavsharma/Desktop/Data Science/midterm.csv")
#3
data %<>%
  rename_with(tolower)

data %>%
  glimpse()
```

```{r}
#4 & 5
data %>%
  count(id,country) %>%
  filter(n>1)
data %>%
  count(country) %>%
  filter(n>1)
```

```{r}
#6
data_tidy <- data %>%
  pivot_longer(!c(id,country), names_to="quarter_num", values_to = "val")
data_tidy %>%
  glimpse()

#7
data_tidy %<>%
  select(id,country,quarter_num,val) %>%
  arrange(id,country)
```

## Iteration 2
### Phase 1
1. Separate columns if needed
2. Check for outliars, set to NA
2. Check for NA values, drop all NA values
4. Recode character columns as factors

```{r}
#1
data_tidy %<>% separate(quarter_num, c("quarter","quarter_num"), sep = "_") %>%
  mutate(quarter_num = as.numeric(quarter_num))

data_tidy %>%
  slice(1:10) %>%
  kable()
```
```{r}
data_tidy %>%
  summary() #No NA
```
##The summary tells us,
1. quarter_num is good
2. val has outliars -999.9990.
3. There are no 'NA' values

```{r}
data_tidy %<>% mutate(
  val = na_if(val, -999.9990)
)
##check 
data_tidy %>%
  pull(val) %>%
  summary()
```
```{r}
data_tidy %<>%
  drop_na()

data_tidy %>%
  summary()
```
```{r}
data_tidy %>%
  glimpse()
```
```{r}
data_tidy %<>%
  mutate(id = as.factor(id), country = as.factor(country), quarter = as.factor(quarter))

data_tidy %>%
  slice(1:20) %>%
  kable()
```

\newpage
#2a
```{r}
data_tidy %>%
  group_by(country, quarter_num) %>%
  summarise(mean = mean(val))
```
#2bcd
```{r}
data2 <- data_tidy %>%
  mutate(dif = abs(mean(val) - median(val))) %>%
  arrange(desc(dif)) %>%
  filter(dif > 0.3)

data2 %>%
  glimpse()
```
###e.
There are no combinations of dif > 0.3

\newpage
#Bonus question
```{r}
data_untidy <- data_tidy %>%
  pivot_wider(!c(quarter,quarter_num,), names_from = country, values_from = val) %>%
  slice(21:28)

data_untidy %>%
  glimpse()
```

